### 2017-11-13 Draft2/Appending.Dynamic.tex
...
Was close...

It would be helpful if the ratio $\dfrac{\FuncTimeDynamicArray(n)}{\FuncTimeGrowthArray(n)}$ could be determined, where $\FuncTimeDynamicArray(n)$ and $\FuncTimeGrowthArray(n)$ denote the average time needed to append $n$ items for dynamic and growth arrays, respectively. However, there is no way to quantify these functions without empirical measurements. Thus, I will approximate this ratio through a different technique: Suppose $n$ elements are appended to an empty collection. How many times is an element stored in an array? I will term the answer to this question the \textbf{write cost} of $n$ appends, and denote it $\FuncWrites(n)$.

@@

I can build on this inequality to receive asymptotic bounds for $\FuncWrites(n)$:
\begin{align*}
\frac n \VarInitCapacity &\FluteLeq \VarGrowthFactor^\VarUseful \FluteLess \frac {\VarGrowthFactor{n}} \VarInitCapacity\\
\frac n \VarInitCapacity &\FluteLeq \VarGrowthFactor^\VarUseful - 1 \FluteLess \frac {\VarGrowthFactor{n}} \VarInitCapacity\\
\frac n {\VarGrowthFactor - 1} &\FluteLeq \left( \frac{\VarGrowthFactor^\VarUseful - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity \FluteLess \frac {\VarGrowthFactor{n}} {\VarGrowthFactor - 1}\\
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n &\FluteLeq \FuncWrites(n) \FluteLess \left( \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1} \right) n
\end{align*}
\HdrSpaceComplex

>>>>

It would be helpful if the ratio $\dfrac{\FuncTimeDynamicArray(n)}{\FuncTimeGrowthArray(n)}$ could be determined, where $\FuncTimeDynamicArray(n)$ and $\FuncTimeGrowthArray(n)$ denote the average time needed to append $n$ items for dynamic and growth arrays, respectively. This represents how much faster growth arrays are than dynamic ones. However, there is no way to quantify these functions without empirical measurements. Thus, I will approximate this ratio through a different technique: Suppose $n$ elements are appended to an empty collection. How many times is an element stored in an array? I will term the answer to this question the \textbf{write cost} of $n$ appends, and denote it $\FuncWrites(n)$.

@@

I can build on this inequality to receive asymptotic bounds for $\FuncWrites(n)$:
\begin{align*}
\frac n \VarInitCapacity &\FluteLeq \VarGrowthFactor^\VarUseful \FluteLess \frac {\VarGrowthFactor{n}} \VarInitCapacity\\
\frac n \VarInitCapacity &\FluteLeq \VarGrowthFactor^\VarUseful - 1 \FluteLess \frac {\VarGrowthFactor{n}} \VarInitCapacity &&\text{(Corollary \ref{coro:PlusConstant}, Theorem \ref{thm:InterchangeableInInequality}, Corollary \ref{coro:InterchangeableInNonStrictInequality})}\\
\frac n {\VarGrowthFactor - 1} &\FluteLeq \left( \frac{\VarGrowthFactor^\VarUseful - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity \FluteLess \frac {\VarGrowthFactor{n}} {\VarGrowthFactor - 1} &&\text{(Theorem \ref{thm:BothSidesInequality}, Corollary \ref{coro:BothSidesNonStrictInequality})}\\
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n &\FluteLeq \FuncWrites(n) \FluteLess \left( \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1} \right) n % Bookmark
\end{align*}
\HdrSpaceComplex

### 2017-11-13 Draft2/Appendix.tex
Was close...

	The statement is much more easily proven for multiplication. Since both $\ExprNToInfty \dfrac{f}{\Tweak{f}}$ and $\ExprNToInfty \dfrac{g}{\Tweak{g}}$ exist,
	\begin{align*}
	\ExprNToInfty \frac{fg}{\Tweak{f}\Tweak{g}} = \left( \ExprNToInfty \frac{f}{\Tweak{f}} \right) \left( \ExprNToInfty \frac{g}{\Tweak{g}} \right) = 1 \cdot 1 = 1
	\end{align*}
	It follows that $fg \sim \Tweak{f}\Tweak{g}$. If the limit for $g$ is flipped before multiplying, the following results:
	\begin{align*}
	\ExprNToInfty \frac{f / g}{\Tweak{f} / \Tweak{g}} = 1
	\end{align*}
	This implies that $\dfrac f g \sim \dfrac {\Tweak{f}} {\Tweak{g}}$.
\end{proof}

\subsubsection{$\sim$ Relations can be Algebraically Manipulated}

@@

\subsection{Asymptotic Functions may be Swapped in Strict Asymptotic Inequalities}

@@

\subsubsection{Asymptotic Functions may be Swapped in Non-Strict Asymptotic Inequalities}

\begin{proof}[Proof (Corollary \ref{coro:InterchangeableInNonStrictInequality})]
	Suppose $f \sim \Tweak{f}$ and $\Tweak{f} \FluteLeq \TweakSecond{f}$. By definition, $\Tweak{f} \FluteLess \TweakSecond{f}$ or $\Tweak{f} \sim \TweakSecond{f}$. In the first case, from Theorem \ref{thm:InterchangeableInInequality} $f \FluteLess \TweakSecond{f}$. In the second case, since $\sim$ is transitive $f \sim \TweakSecond{f}$. In both cases, it is true that $f \FluteLeq \TweakSecond{f}$.

	The second statement is, again, proved in a similar manner. If $\Tweak{g} \FluteLeq \TweakSecond{g}$, then $\Tweak{g} \FluteLess \TweakSecond{g}$ or $\Tweak{g} \sim \TweakSecond{g}$. In the first case, $\Tweak{g} \FluteLess g$; in the second, $\Tweak{g} \sim g$. In both cases, $\Tweak{g} \FluteLeq g$.
\end{proof}

\subsubsection{$\FluteLess$ and $\FluteLeq$ Relations can be Algebraically Manipulated}

\begin{proof}[Proof (Theorem \ref{thm:BothSidesInequality})]
	Suppose that $f \FluteLess g$. By definition, $\ExprNToInfty \dfrac{f}{g} < 1$. For any function $h$, $\ExprNToInfty \dfrac{fh}{gh} < 1$ and $\ExprNToInfty \dfrac{f / h}{g / h} < 1$, so respectively $fh \FluteLess gh$ and $\dfrac{f}{h} \FluteLess \dfrac{g}{h}$.

	If I suppose $f \FluteLeq g$, then $f \FluteLess g$ or $f \sim g$. The first statement implies $fh \FluteLess gh$, and the second implies $fh \sim gh$ (from Corollary \ref{coro:BothSides}). In either case, $fh \FluteLeq gh$. Using the same logic, it is true that $\dfrac{f}{h} \FluteLeq \dfrac{g}{h}$.
\end{proof}

>>>>

	The statement is proven much more easily for multiplication. Since both $\ExprNToInfty \dfrac{f}{\Tweak{f}}$ and $\ExprNToInfty \dfrac{g}{\Tweak{g}}$ exist,
	\begin{align*}
	\ExprNToInfty \frac{fg}{\Tweak{f}\Tweak{g}} = \left( \ExprNToInfty \frac{f}{\Tweak{f}} \right) \left( \ExprNToInfty \frac{g}{\Tweak{g}} \right) = 1 \cdot 1 = 1
	\end{align*}
	It follows that $fg \sim \Tweak{f}\Tweak{g}$. If the limit for $g$ is flipped before multiplying, the following results:
	\begin{align*}
	\ExprNToInfty \frac{f / g}{\Tweak{f} / \Tweak{g}} = 1
	\end{align*}
	This implies that $\dfrac f g \sim \dfrac {\Tweak{f}} {\Tweak{g}}$.
\end{proof}

\subsubsection{$\sim$ Equations can be Algebraically Manipulated}

@@

\subsection{Asymptotic Functions may be Interchanged in $\FluteLess$ Inequalities}

@@

\subsubsection{Asymptotic Functions may be Interchanged in $\FluteLeq$ Inequalities}

\begin{proof}[Proof (Corollary \ref{coro:InterchangeableInNonStrictInequality})]
	Suppose $f \sim \Tweak{f}$ and $\Tweak{f} \FluteLeq \TweakSecond{f}$. By definition, $\Tweak{f} \FluteLess \TweakSecond{f}$ or $\Tweak{f} \sim \TweakSecond{f}$. In the first case, from Theorem \ref{thm:InterchangeableInInequality}, $f \FluteLess \TweakSecond{f}$. In the second case, since $\sim$ is transitive, $f \sim \TweakSecond{f}$. In both cases, it is true that $f \FluteLeq \TweakSecond{f}$.

	The second statement is proved similarly. It is given that $g \sim \TweakSecond{g}$ and $\Tweak{g} \FluteLeq \TweakSecond{g}$; the latter implies that either $\Tweak{g} \FluteLess \TweakSecond{g}$ or $\Tweak{g} \sim \TweakSecond{g}$. In the first case, $\Tweak{g} \FluteLess g$; in the second, $\Tweak{g} \sim g$. In both cases, $\Tweak{g} \FluteLeq g$.
\end{proof}

\subsection{$\FluteLess$ Inequalities can be Algebraically Manipulated}

\begin{proof}[Proof (Theorem \ref{thm:BothSidesInequality})]
	Suppose that $f \FluteLess g$. By definition, $\ExprNToInfty \dfrac{f}{g} < 1$. For any function $h$, $\ExprNToInfty \dfrac{fh}{gh} < 1$ and $\ExprNToInfty \dfrac{f / h}{g / h} < 1$, so respectively $fh \FluteLess gh$ and $\dfrac{f}{h} \FluteLess \dfrac{g}{h}$.
\end{proof}

\subsubsection{$\FluteLeq$ Inequalities can be Algebraically Manipulated}

\begin{proof}[Proof (Corollary \ref{coro:BothSidesNonStrictInequality})]
	If I suppose $f \FluteLeq g$, then $f \FluteLess g$ or $f \sim g$. The first case implies $fh \FluteLess gh$ from Theorem \ref{thm:BothSidesInequality}, and the second implies $fh \sim gh$ from Corollary \ref{coro:BothSides}. In either case, $fh \FluteLeq gh$. Using the same logic, it is true that $\dfrac{f}{h} \FluteLeq \dfrac{g}{h}$.
\end{proof}

\subsection{Same-Order Terms may be Added to Both Sides of $\FluteLess$ Inequalities}

\begin{proof}[Proof (Theorem \ref{thm:BothSidesInequalityAdd})]
	From Theorem \ref{thm:InterchangeableInInequality}, we may substitute $f$ and $g$ in the inequality with $c_1 h$ and $c_2 h$ to receive $c_1 h \FluteLess c_2 h$. Then $\ExprNToInfty \dfrac{c_1 h}{c_2 h} = \dfrac{c_1}{c_2} < 1$, so $c_1 < c_2$.

	Now, consider that $c_1 h + h \sim f + h$ and $g + h \sim c_2 h + h$ from Corollary \ref{coro:BothSides}. By definition, the limits $\ExprNToInfty \dfrac{(c_1 + 1) h}{f + h}$ and $\ExprNToInfty \dfrac{g + h}{(c_2 + 1) h}$ exist and are equal to $1$. Then,
	\begin{align*}
	\ExprNToInfty \frac{f + h}{g + h} &= \left( \ExprNToInfty \frac{f + h}{g + h} \right) \left( \ExprNToInfty \dfrac{(c_1 + 1) h}{f + h} \right) \left( \ExprNToInfty \dfrac{g + h}{(c_2 + 1) h} \right)\\
	&= \ExprNToInfty \frac{(c_1 + 1) h}{(c_2 + 1) h}\\
	&= \frac{c_1 + 1}{c_2 + 1}
	\end{align*}
	Since $c_1 < c_2$, $c_1 + 1 < c_2 + 1$ so $\dfrac{c_1 + 1}{c_2 + 1} < 1$. Thus, $f + h \FluteLess g + h$.
\end{proof}

\subsubsection{Same-Order Terms may be Added to Both Sides of $\FluteLeq$ Inequalities}

\begin{proof}[Proof (Corollary \ref{coro:BothSidesNonStrictInequalityAdd})]
	Given $f \FluteLeq g$, it is true that either $f \FluteLess g$ or $f \sim g$. In the first case, from Theorem \ref{thm:BothSidesInequalityAdd} it follows that $f + h \FluteLess g + h$. In the second case, from Theorem \ref{coro:BothSides} it follows that $f + h \sim g + h$. In both cases, $f + h \FluteLeq g + h$.
\end{proof}

### 2017-11-13 Draft2/AsymptoticNotation.tex
Was close...

I define various properties of the $\sim$ relation here, which later proofs will use. The properties themselves are proved in the appendix.

In the following theorems and their proofs, variables with the letters $f$, $g$, or $h$ denote arbitrary functions that are positive and non-decreasing for sufficiently large $n$.

The following theorem states that $\sim$ meets the requirements of an equivalence relation.

@@

The following theorem states that if two functions are asymptotic, then they belong to the same big-O class.

@@

The following theorem states that if two functions are asymptotic, then they can be interchanged in an asymptotic inequality.

@@

\begin{corollary}
\label{coro:InterchangeableInNonStrictInequality}
	This theorem also holds for $\FluteLeq$: If $f \sim \Tweak{f}$ and $\Tweak{f} \FluteLeq \TweakSecond{f}$, then $f \FluteLeq \TweakSecond{f}$. If $g \sim \TweakSecond{g}$ and $\Tweak{g} \FluteLeq \TweakSecond{g}$, then $\Tweak{g} \FluteLeq g$.
\end{corollary}

The following theorem states that asymptotic inequalities can be algebraically manipulated, in the same vein as Corollary \ref{coro:BothSides}. However, they do not always hold if the same quantity is added to both sides; for example, $1 \FluteLess 2$, but $n + 1 \sim n + 2$.

\begin{theorem}
\label{thm:BothSidesInequality}
	A $\FluteLess$ or $\FluteLeq$ inequality holds if both sides are multiplied or divided by the same quantity.
\end{theorem}

>>>>

I define various properties of the $\sim$ relation here, which proofs in later sections will use. The properties themselves are proved in the appendix.

In the following theorems and their proofs, variables with the letters $f$, $g$, or $h$ denote functions that, for sufficiently large $n$, are positive and non-decreasing.

The following theorem states that $\sim$ is a valid equivalence relation.

@@

The following theorem states that asymptotic functions belong to the same big-O class.

@@

The following theorem states that asymptotic functions can be interchanged in an asymptotic inequality.

@@

\begin{corollary}
\label{coro:InterchangeableInNonStrictInequality}
	This theorem also holds for $\FluteLeq$.
\end{corollary}

The following theorem states that asymptotic inequalities can be algebraically manipulated, in the same vein as Corollary \ref{coro:BothSides}.

\begin{theorem}
\label{thm:BothSidesInequality}
	A $\FluteLess$ inequality holds if both sides are multiplied or divided by the same quantity.
\end{theorem}

\begin{corollary}
\label{coro:BothSidesNonStrictInequality}
	This theorem also holds for $\FluteLeq$.
\end{corollary}

Notice that the above theorem does not include addition. Asymptotic inequalities do not always hold if the same quantity is added to both sides; for example, $1 \FluteLess 2$, but $n + 1 \sim n + 2$. In order for them to hold, certain conditions must be met.

\begin{theorem}
\label{thm:BothSidesInequalityAdd}
	Suppose $f \FluteLess g$. If $f \sim c_1 h$ and $g \sim c_2 h$ for some $c_1, c_2 > 0$, then $f + h \FluteLess g + h$.
\end{theorem}

\begin{corollary}
\label{coro:BothSidesNonStrictInequalityAdd}
	This theorem also holds for $\FluteLeq$.
\end{corollary}
